The Local Group in the Cosmic Web

### 2014-10-21 paper/referee_reply.tex

\item
{\bf One of the main aspects of the paper is the alignment between the three eigenvectors of the tidal tensor and the two vectors defined by Local Group analogues (the normal to their orbital plane and the line connecting the two). Figure 3 and Figure 4 show this (variants are shown in figure 5 \& 6 as well). What these figures lack is an estimation - a quantification - of the statistical significance of the result. This can be easily done by constructing 3xN "sister" samples with the same number of elements as the 2sigma, 3sigma and general samples, uniformly drawn in the interval [0,1] and comparing them. For example in the case of figure 3 \& 4 one could imagine plotting the 2 sigma spread in the cumulative distribution of such random distributions. In Fig 5 some kind of spread indicating the variance in the median computations could be included (along with the medians and spread for a uniform distribution). In Figure 6 the authors could also include an expectation from random distributions.}

We have followed the advice from the referee. Now Figures 3, 4, 5 and 6 include a shadowed region quantifying the expectations from a random distribution. The captions include an explanation on how the numbers were estimated.

Our estimation proceeds as follows. We construct a sample of 3D vectors with the same number of elemets as the 2sigma sample. We use the 2sigma sample to show the upper bound on the uncertainties.

These vectors have a random distributions. This allows us to build the integrated $\mu$ distributions. We repeat the exercise 10000 times to compute the median together with the $5\%$ and $95\%$ percentiles for the quantities of interest. In the case of Figures 3,4 and 6 it corresponds to the fraction of elements with values $<\mu$; in the case of Figure 5 it corresponds to the median value of $\mu$.

@@

\item
{\bf My other significant point is the smoothing scale. The study examines the web on a fixed scale namely ~1.4Mpc. However since the scale is fixed it measures different aspects of the environment for different LG analogues. This is explicitely shown in Fig2 left. A halo of mass 5e13Mdot could have a virial radius of upto 1 or 1.3Mpc. For these haloes the smoothing is basically on the scale of Rvir. Whereas for haloes with masses of 5e11 Mdot the virial radius is probably less than 200kpc and your 1.4Mpc smoothing is equivalent to 7 times the virial radius.  Although I am quite sure that the tidal tensor eigenvectors are stable across these different scales, I'm not quite sure the web classification, web ellipticity or web prolateness are. Perhaps you should convince the reader of this issue.}

We tested the robustness of our results with a smaller smoothing scale (0.7 Mpc, from the $512^3$ grid data available on the Multidark Database. This is mentioned in the paper.

Concerning a larger smoothing scale we base the robustness on already published results. In Forero-Romero et al. 2009  we presented a through study of the environment as defined by the Tweb in terms of the two free parameters of the model: $\lambda_{th}$ and the smoothing scale. In that work we changed the smothing scale from 0.5 Mpc/h up to 12.5 Mpc/h and quantified the changes in the volume and mass filling fractions (VFF and MFF) for each web environment. We found that for the values of $\lambda_{th}\approx 2$ the VFF and MFF only change on the order of $5\%$. In the thorough paper by Cautun et al. 2014 they also show that the web properties are insensitive to changes of factor of $2$ on the smoothing scale.

We have added a short paragraph reminding the reader about these points.

@@

We have extended the mass range to 13.5 in the plots.

\item{ In the discussion the authors say that "the MW satellites are located at high galactic latitude, almost perpendicular to the direction of M31." I take issue wth the "almost perpendicular" comment. The plane of MW satellites is between 32 and 45 degrees from the direction to M31 depending on how it is defined. This is a small point of contention in the literature because people define it differently given the uncertainties and incompleteness limits that drive the fact that most of these galaxies are found at high galactic latitude.

Using a fairly conservative definition (ie the 11 brightest dwarfs for which we have fairly good completeness limits) Pawlowski et al 2012 and Shaya \& Tully 2014 both find something between 32 and 45 degrees. If the authors want to stick to the "almost perpendicular" statement they will need to justify it by citation and explanation}

We have included the exact values in the Discussion section.

### 2014-10-21 paper/main.tex
full pass over the manuscript

Another feature of the Local Group is the relatively low velocity
dispersion of nearby galaxies up to $\sim 8$ Mpc \citep[][and
  references therein]{1975ApJ...196..313S,2011MNRAS.415L..16A}. 
The environment around the Local Group has density quite close to the
average density of the universe
\citep{2003ApJ...596...19K,2005AJ....129..178K}. 
In addition, the closest massive galaxy cluster, the Virgo Cluster, is
$\approx 16.5\ \mpc$ away \citep{2007ApJ...655..144M}.

@@

\subsection{The Bolshoi simulation}
We use the Bolshoi simulation of $\Lambda$CDM cosmology: $\Omega_{\rm
  m}=1-\Omega_{\Lambda}=0.27$, $H_0=70\,\rm km/s/Mpc$,
$\sigma_8=0.82$, $n_s=0.95$ \citep{2011ApJ...740..102K}, compatible
with the constraints from the WMAP satellite
\citep{hinshaw_etal13}. The simulation followed the evolution of dark
matter in a $250 \hmpc$ box with spatial resolution of $\approx
1h^{-1}$~kpc and mass resolution of $m_{\rm p}=1.35\times 10^8\ \rm
M_{\odot}$. Halos are identified with the BDM algorithm
\citep{1997astro.ph.12217K}. The BDM algorithm is  a spherical
overdensity halo finding algorithm and is designed to identify both
host halos and subhalos.

@@

@@

To construct a sample of the MW-M31 pairs at $z\approx 0$, we use a
series of simulation snapshots  at $z<0.1$ (since the last $\approx
1.3$ Gyr) spaced by $\approx 150-250$ Myr. 
This is done because a particular configuration of MW and M31 is transient and
corresponds to a relatively small number of systems at one
snapshot. 
By using multiple snapshots we can increase the sample of systems in
such configuration during a period of time in which secular
cosmological evolution is small.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{fig5.pdf}
\caption{Mass dependency of the median value for the dot product
  between the normal vector $\hat{n}$ and each one of the
  eigenvectors.  The lines show the trends for the general sample.
The error bars in the 2$\sigma$ and 3$\sigma$ points
  correspond to jackknife estimates. 
 The shaded region shows the expectation for a random distribution
 without any preferential alignment, it encloses the $5\%$ and $95\%$
 percentiles of 5000  flat distributions for $\mu$ generated with the
 same number of points as the $2\sigma$ sample.   
\label{fig:median_alignment_n}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{fig6.pdf}
\end{center}
\caption{Alignment between the two angular momentum vectors of the two
  halos in the pair.
  The shaded region shows the expectation for a random
  distribution without a preferential alignment as it is described in
  Figure \ref{fig:median_alignment_n}.
    \label{fig:jj_alignment}}  
\end{figure}

@@

Can the increase of the alignment strength in the different samples be
explained as a consequence of different mass dependencies as in the
results in the previous section? The answer is no. 
In Figure \ref{fig:median_alignment_n} we show the median of
$\mu\equiv\hat{e}_i\cdot\hat{n}$ in different mass bins.
Lines show the median $\mu$-mass relation for the three 
eigenvectors in the GS, dots represent the results for the
2$\sigma$-3$\sigma$ samples. 
Around the total mass $1.5-2.0 \times 10^{12}$ \msun in the RS
samples, the median of $\mu$ decreases as we tighten constraints
on the pairs, but not as a result of the narrowing mass selection.

@@

Figure \ref{fig:jj_alignment} shows the cumulative distribution of dot
product between angular momentum of the  two halos. We find a a slight
alignment of spin vectors for the $2\sigma$ sample with a median around
$|\mu|\sim 0.7$ ($45$ degrees) which is barely above the no-alignment expectation marked by the shaded region. Furthermore, we found no significant
alignment with the pair orbital angular momentum nor the cosmic web. This
absense of alignment consistent with recent results that show that the spin
alignment of halos of masses around $10^{12}$\Msun is very weak
\citep{ForeroRomero2014}.

@@

How can we relate these alignments to the observed LG? 
To evaluate this point we use observational information for the
satellite distribution around the LG galaxies.  The MW satellites are
located at high galactic latitudes forming a planar structure that
forms an angle of $42^{\circ}-52^{\circ}$, $\mu=0.6-0.7$ with the
vector joining the MW and M31; while the M31 satellites are on a plane
that lies on the same vector. \citep{Pawloski2013,Shaya2013}.

@@

Taking this three points together (the alignment of MW-M31 along
$\hat{e}_3$, the observed satellite alignments and the preferential
infall along $\hat{e}_3$) we have that the M31 plane of satellites is
completely consistent with the average alignment picture we have
described. However, the planes defined by the MW satellites should
also be close to parallel with respect to the vector joining MW and
M31 and the plane of M31 satellites. Instead, this plane has
$\mu=0.6-0.7$ with respect to these to directions.

This apparent contradiction can be solved if one considers that the
subhalo infall properties depend on the environment at {\it the
  merger time}, while the alignments for yet unvirialized pairs such as the LG
depend on the current state of the cosmic web. This also explain the absence
of a strong alignment with halo spin, because it reflects past
configurations of the web, not the current one. From this perspective,
a joint consideration of the alignments between the dominant
halos in the LG and their satellites, which also feature strong
signals \citep{2005A&A...431..517,Pawloski2013,Shaya2013}, should
inform us about the structural evolution of the cosmic web around the
LG.

Our results also raise the need to observationally constraint the
alignments of LG pairs with their cosmic web environment. To this end one could use mass reconstructions from large surveys
\citep{2009MNRAS.394..398W,2011MNRAS.417.1303M,2014arXiv1407.3451W,2014arXiv1406.1004N}
or filament finders \citep{2010MNRAS.407.1449G,2011MNRAS.414..350S}
to select a LG sample to quantify the alignments with the filamentary
structure. 
This allows not only a direct quantification of how common are the LG
alignments but also provides a new test of $\Lambda$CDM.

>>>>

Another feature of the Local Group is the relatively low velocity
dispersion of nearby galaxies up to $\sim 8$ Mpc \citep[][and
  references therein]{1975ApJ...196..313S,2011MNRAS.415L..16A}. 
The environment around the Local Group has density quite close to the
average density of the universe
\citep{2003ApJ...596...19K,2005AJ....129..178K}. 
In addition, the closest massive galaxy cluster, the Virgo Cluster, is
$\sim 16.5\ \mpc$ away \citep{2007ApJ...655..144M}.

@@

\subsection{The Bolshoi simulation}
We use the Bolshoi simulation of $\Lambda$CDM cosmology: $\Omega_{\rm
  m}=1-\Omega_{\Lambda}=0.27$, $H_0=70\,\rm km/s/Mpc$,
$\sigma_8=0.82$, $n_s=0.95$ \citep{2011ApJ...740..102K}, compatible
with the constraints from the WMAP satellite
\citep{hinshaw_etal13}. The simulation followed the evolution of dark
matter in a $250 \hmpc$ box with spatial resolution of $\sim
1h^{-1}$ kpc and mass resolution of $m_{\rm p}=1.35\times 10^8\ \rm
M_{\odot}$. Halos are identified with the BDM algorithm
\citep{1997astro.ph.12217K}. The BDM algorithm is  a spherical
overdensity halo finding algorithm and is designed to identify both
host halos and subhalos.

@@

We have
veryfied that the main trends reported in this paper remain unchanged
if we use the results from a $512^3$ grid smoothed over scales of
$s=0.48$\hMpc. Small changes of a factor of $\sim2$ in the smoothing
scale do not significantly impact the cosmic web as it 
was shown in \cite{Tweb} from the study of the volume and mass
filling fractions for different smoothing scales. Other works have
also shown that broad features the cosmic web, as defined in this
paper, are robust to factor of $\sim2$ changes on the scale defining the web
\citep{Cautun2014}.

@@

To construct a sample of the MW-M31 pairs at $z\sim 0$, we use a
series of simulation snapshots  at $z<0.1$ (since the last $\sim
1.3$ Gyr) spaced by $\sim 150-250$ Myr. 
This is done because a particular configuration of MW and M31 is transient and
corresponds to a relatively small number of systems at one
snapshot. 
By using multiple snapshots we can increase the sample of systems in
such configuration during a period of time in which secular
cosmological evolution is small.

@@

\begin{figure*}
\begin{center}
  \includegraphics[width=0.45\textwidth]{fig5a.pdf}
  \includegraphics[width=0.45\textwidth]{fig5b.pdf}
\caption{Mass dependency of the median value for the dot product
  between the normal/radial vector $\hat{n}$/$\hat{r}$ (left/right)
  and each one of the  eigenvectors.  The lines show the trends for
  the general sample. The error bars in the 2$\sigma$ and 3$\sigma$ points
  correspond to jackknife estimates. 
  The shaded region shows the expectation for a random distribution
 without any preferential alignment, it encloses the $5\%$ and $95\%$
 percentiles of 5000  flat distributions for $\mu$ generated with the
 same number of points as the $2\sigma$ sample.   
\label{fig:median_alignment_n}}
\end{center}
\end{figure*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.45\textwidth]{fig6.pdf}
\end{center}
\caption{Alignment between the two angular momentum vectors of the two
  halos in the pair.
  The shaded region shows the expectation for a random
  distribution without a preferential alignment as it is described in
  Figure \ref{fig:median_alignment_n}.
    \label{fig:jj_alignment}}  
\end{figure}

@@

Can the increase of the alignment strength in the different samples be
explained as a consequence of different mass dependencies as in the
results in the previous section? The answer is no. 
In Figure \ref{fig:median_alignment_n} we show the median of
$\mu\equiv\hat{e}_i\cdot\hat{n}$ and $\mu\equiv \hat{e}_i\cdot\hat{r}$
in different mass bins. 
Lines show the median $\mu$-mass relation for the three
eigenvectors in the GS, dots represent the results for the
2$\sigma$-3$\sigma$ samples.  Around the total mass $1.5-2.0 \times
10^{12}$ \msun in the RS samples, the median of $\mu$ clearly from the
2$\sigma$ and 3$\sigma$ samples clearly differs from the GS results.
Although in the GS the alignments have slight changes with mass, this
is not enough to explain the results for the reduced samples.

@@

Figure \ref{fig:jj_alignment} shows the cumulative distribution of dot
product between angular momentum of the  two halos. We find a a slight
alignment of spin vectors for the $2\sigma$ sample with a median around
$|\mu|\sim 0.7$ ($45$ degrees) which is barely above the no-alignment expectation marked by the shaded region. Furthermore, we found no significant
alignment with the pair orbital angular momentum nor the cosmic web.

The absence of alignment for the spins is consistent with different studies of
spin alingnemnt that mark the range $10^{12}$\Msun as either a transition mass
from alignment into anti-alignment or from no-alignment into alignment
\citep{Hahn2007,AragonCalvo2007,Zhang2009,Codis2012,Trowland2013,Libeskind2013,ForeroRomero2014}. We refer the reader to Table 2 in \cite{ForeroRomero2014}
for a review on the different results of spin alignment with the
cosmic web in cosmological simulations. It seems that this transition
has to do with the changes on the major merger rate as a dark matter
halo approaches the $10^{12}$\Msun mass \citep{2009MNRAS.399..762F,Codis2012}.

@@

How can we relate these alignments to the observed LG? 
To evaluate this point we use observational information for the
satellite distribution around the LG galaxies.  The MW satellites are
located at high galactic latitudes forming a planar structure that
forms an angle of $42^{\circ}-52^{\circ}$, $\mu=0.6-0.7$ with the
vector joining the MW and M31; while the M31 satellites are on a plane
that lies on the same vector \citep{Pawlowski2013,Shaya2013}.

@@

Taking this three points together (the alignment of MW-M31 along
$\hat{e}_3$, the observed satellite alignments and the preferential
infall along $\hat{e}_3$) we have that the M31 plane of satellites is
completely consistent with the average alignment picture we have
described. However, the plane defined by the MW satellites should
also be close to parallel with respect to the vector joining MW and
M31 and the plane of M31 satellites. Instead, this plane has
$\mu=0.6-0.7$ with respect to these directions. The spatial location
of MW satellites raise a potential contradiction with the average expectations
from $\Lambda$CDM, a point that has been raised before
\citep{Pawlowski2012,Pawlowski2014}.

This apparent contradiction can be solved if one considers that the
subhalo infall properties depend on the environment at {\it the
  merger time}, while the alignments for yet unvirialized pairs such
as the dominant galaxis in the LG depend on the current state of the
cosmic web. This also explain the absence of a strong alignment with
halo spin, because it reflects past configurations of the web, not the
current one. From this perspective, a joint consideration of the
alignments between the dominant halos in the LG and their satellites,
which also feature strong signals
\citep{2005A&A...431..517K,Pawlowski2013,Shaya2013}, should inform us
about the structural evolution of the cosmic web around the LG.

Our results raise the need to observationally constrain the
alignments of LG pairs with their cosmic web environment. To this end
one could use mass reconstructions from large surveys
\citep{2009MNRAS.394..398W,2011MNRAS.417.1303M,2014arXiv1407.3451W,2014arXiv1406.1004N}
or filament finders \citep{2010MNRAS.407.1449G,2011MNRAS.414..350S} to
select a LG sample to quantify the alignments with the surrounding
filamentary structure. This would allow not only a direct
quantification of how common are the LG alignments but also provide a
new test of $\Lambda$CDM.  We thank the referee Noam Libeskind for
suggestions that helped us to clarify and strenghten the results
presented in the paper.
